Identifiyng Inter-Genomic Repeats Using Fast, Approximate Betweenness Centrality

### 2015-05-04 manuscript/manuscript.tex
evaluation script added
modified report

% first the title is needed
\title{Lecture Notes in Computer Science:\\Authors' Instructions
for the Preparation\\of Camera-Ready
Contributions\\to LNCS/LNAI/LNBI Proceedings}

@@

\subsection*{Networkit}
We investigated the performance of approximate betweenness centrality algorithms on the metagenomic graphs. We used the implementation in Networkit\cite{networkit} library. Its a c++ library with parallel implementation of many graph algorithms. We used the implementation of approximate betweenness centrality algorithm by\cite{matteo}. This algorithm is implemented in parallel.We also investigated the performance of algorithm by (cite other) on graphs. This algorithm is not parallelized. It takes number of sample nodes as an input.

@@

>>>>

% first the title is needed
\title{Identifiyng Inter-Genomic Repeats Using Fast, Approximate Betweenness Centrality}

@@

\subsection*{Networkit}
We investigated the performance of approximate betweenness centrality algorithms on the metagenomic graphs to find inter-genome repeats. We used the implementation in Networkit\cite{networkit} library. It provides implementation for various graph algorithms. We are particularly interested in the implementation of betweenness centrality algorithms. We make use of two approximate betweenness centrality algorithms provided by NetworKit.

NetworKit provides fast parallel implementation of approximate betweenness centrality algorithm by Matteo et al. \cite{matteo}. The implementation first approximates diameter of of graph. After this, it calculates the number of shortest paths to be sampled. It processes shortest paths in the sample in parallel in multithreaded fashion and then aggregates the centrality scores for each shortest paths once every thread finishes processing of shortest path assigned to it.

We also investigate the performance of algorithm by Geisberger et al. \cite{sanders}. NetworKit provides single threaded implementation of this algorithm. It first creates a sample of $n$ nodes at random from a given graph. It then runs single source shortest path algorithm for each of those nodes. It calculates the betweenness scores for each node by the modified aggregation scheme proposed in the algorithm.

Here are few questions we want to investigate based on these algorithms:

\begin{itemize}
\item \textbf{Scalability}: How these algorithms scale on large metagenomic repeats graph? Does parallelized algorithm provides quick results or taking sample of fewer nodes provides quick results?
\item \textbf{Accuracy}: Which of these algorithms provides better accuracy while determining inter-genomic repeats?
\item \textbf{Sample Size}: For the algorithm by Geisberger et al. \cite{sanders}, how should a sample size be defined so that it would produce fast and accurate results? We try to come up with a heuristic to define sample size
\end{itemize}

@@

\section{Results} 
We evaluated algorithms for betweenness centrality to find inter-genome repeats on real datasets described above. We use the notion of sensitivity and specificity to define the accuracy of repeat detection. We define following terms:
\begin{itemize}
\item \textbf{True positives:} Number of nodes marked as repeat that is found within at least 2 genomes
\item \textbf{False positives:} Number of nodes marked as repeat that is only repetitive within a single genome
\item \textbf{True negatives:} Number of nodes marked as non-repeat that is only repetitive within a single genome
\item \textbf{False negatives:} Number of nodes marked as non-repeat that is found within at least 2 genomes
\end{itemize}

A good repeat detection should have both good sensitivity and specificity. Sensitivity reflects how many true repeats are detected. Detecting to few repeats can lead to assembly errors in scaffolding. Specificity reflects the false positives. Detecting too many repeats leads to a suboptimal assembly as these contigs do not fully participate in scaffolding.
